Effects of bariatric surger and weight loss on resting-state functional connectivity

### 2020-12-02 intro_rsfmri.Rmd
deleted abbreviations

<!-- Reward network -->
Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population. Thus, one recently emerging research focus is how brain metabolism and functional connectivity changes in response to surgery, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
The reward network comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antIns) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016]. These regions also seem to be altered after bariatric surgery [@Lips_2014; @Hogenkamp_2016; @Garcia_2012]. <!-- Zitationen checken -->
<!-- Normalization of connectivity --> Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions). But these differences seemed to normalize, as trends reversed when looking at the resting state connectivity after compared to before the surgery [@Duan_2020]. Resting state activitiy increased in subcortical regions, such as the insula and striatum<!-- specifically putamen --> [@Wiemerslage_2016]. 
<!-- mechanism reward and appetite, overeating --> Interestingly, patients report changes in food preferences and eating behavior after the surgery that may influence weight loss beyond malabsorption of nutrients and restricted food ingestion. Those psychophysiological and behavioral effects are presumably related to post-surgical changes in visceral sensory perceptions and endocrinological signaling. Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017; @Nudel_2019].
<!-- results in lean -->

@@

The default mdoe network is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (MPFC) and inferior and lateral parietal cortex (IPC, LPC; approx. BA 39) [@Raichle_2015]. The right ventral anterior cingulate cortex (VACC), superior frontal cortex (SFC), and inferior temporal cortex (ITC) show weaker but also reliable association with the PCC over the course of one year. Even broader conceptions count left VACC, middle temporal gyrus (MTG), and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
<!-- study results --> 
Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2017].
Generally, studies show an increased activation in the default mode network, especially in the precuneus and posterior cingulate cortex (PCC) in obese individuals [@Kullmann_2011; @].
Comparing spontaneous activitiy before and after surgery shows a decrease in areas including superior frontal regions<!-- gehört zu SFC -->, and a decrease superior parietal lobule when in sated state [@Wiemerslage_2016].
Concerning spontaneuous functional connectivity in these regions, patient having undergone bariatric surgery resemble normal-weight individuals rather than obese individual in having a weaker connectivity within the default mode network [@Frank_2013].
A longitudinal study investigating confirmed that while there was stronger connectivity in the default model network <!-- (PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG) --> in obese patients before the surgery, it seemed weakened over the course of one year. Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
Consistent with these findings, other studies reported a reduction of functional connectivity of these and other regions of the default mode network, such as the VMPFC <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery [@Li_2018].

<!-- The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated and especially the role of weight and weight change. -->
... 
It has been stated that among patients undergoing surgery, change in BMI was<!-- negatively --> correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
On the other hand, there are also indicators of reduced functional integrity, e.g. for the anterior cingulate cortex (ACC) at resting state compared to lean individuals [@Kullmann_2011]. ...

>>>>

<!-- Reward network -->
Little is known about the potential beneficial effects of bariatric surgery on brain function in the obese population. Thus, one recently emerging research focus is how brain metabolism and functional connectivity changes in response to surgery, in particular with regard to hedonic motivations for food, which play a fundamental role in weight relapse.
The reward network comprising the ventromedial prefrontal cortex (vmPFC), the striatum (STR) but also the anterior insula (antINS) has been linked to food valuation processes in decision-making in lean individuals [@Bartra_2013; @Hare_2011; @Hutcherson_2012; @Schmidt_2018; @Wiemerslage_2016]. These regions also seem to be altered after bariatric surgery [@Lips_2014; @Hogenkamp_2016; @Garcia_2012]. <!-- Zitationen checken -->
<!-- Normalization of connectivity --> Obese individuals exhibited increased connectivity between frontal and mesolimbic regions and weaker connectivity within regions of the mesolimbic system (encompassing limbic and striatal regions). But these differences seemed to normalize, as trends reversed when looking at the resting state connectivity after compared to before the surgery [@Duan_2020]. Resting state activitiy increased in subcortical regions, such as the insula and striatum<!-- specifically putamen --> [@Wiemerslage_2016]. 
<!-- mechanism reward and appetite, overeating --> Interestingly, patients report changes in food preferences and eating behavior after the surgery that may influence weight loss beyond malabsorption of nutrients and restricted food ingestion. Those psychophysiological and behavioral effects are presumably related to post-surgical changes in visceral sensory perceptions and endocrinological signaling. Corresponding information is processed by the fronto-mesolimbic system and affects appetite and reward signalling which may influence weight regulation [@Mulla_2017; @Nudel_2019].
<!-- results in lean -->

@@

The default mdoe network is mainly composed of the posterior cingulate cortex (PCC)/precuneus, the medial prefrontal cortex (mPFC) and inferior and lateral parietal cortex (approx. BA 39) [@Raichle_2015]. The right ventral anterior cingulate cortex (vACC), superior frontal cortex, and inferior temporal cortex show weaker but also reliable association with the PCC over the course of one year. Even broader conceptions count left VACC, middle temporal gyrus, and parahippocampal cortex, although associations with the PCC are not as stable [@Chou_2012].
<!-- study results --> 
Functional connectivity of this network has been shown to be linked to BMI [@Beyer_2017] and altered after bariatric surgery [@Olivo_2017].
Generally, studies show an increased activation in the default mode network, especially in the precuneus and posterior cingulate cortex (PCC) in obese individuals [@Kullmann_2011; @].
Comparing spontaneous activitiy before and after surgery shows a decrease in areas including superior frontal regions<!-- gehört zu SFC -->, and a decrease superior parietal lobule when in sated state [@Wiemerslage_2016].
Concerning spontaneuous functional connectivity in these regions, patient having undergone bariatric surgery resemble normal-weight individuals rather than obese individual in having a weaker connectivity within the default mode network [@Frank_2013].
A longitudinal study investigating confirmed that while there was stronger connectivity in the default model network <!-- (PCC with anterior STG and anterior temporal fusiform gyrus and the Precuneus with the anterior STG) --> in obese patients before the surgery, it seemed weakened over the course of one year. Because differences between lean controls and obese individuals that existed prebariatric surgery were no longer detectable after the surgery, the authors argue for a normalization in functional connectivity<!-- and PCC/PC connectivity with anterior STG are most affected by the surgery -->[@Olivo_2017].
Consistent with these findings, other studies reported a reduction of functional connectivity of these and other regions of the default mode network, such as the vmPFC <!-- with the dorsolateral prefrontal cortex (DLPFC) --> after the surgery [@Li_2018].

<!-- The exact mechanisms of how bariatric surgery affects functional plasticity needs to be investigated and especially the role of weight and weight change. -->
... 
It has been stated that among patients undergoing surgery, change in BMI was<!-- negatively --> correlated with the changes in connectivity of the default mode network, namely between the PCC/precuneus and the DLPFC [@Li_2018].
On the other hand, there are also indicators of reduced functional integrity, e.g. for the ACC at resting state compared to lean individuals [@Kullmann_2011]. ...
